Last glacial cycle glacier erosion potential in the Alps

### 2021-05-27 articles/erosion/response-referee1.md

> Authors decided to keep the paper short, referring most information on
> glacial model set up and resulting ice cover to the Seguinot et 2018 paper.
> This makes it no always very easy to read and needs checking in the original
> manuscript (for example, ice cover deviations model – field bases LGM
> extension).

Indeed, it is our deliberate choice to avoid paraphrasing or duplicating
content from the 2018 paper, where the preparation of the text and figures also
involved a different author team. We believe this is justified given that both
papers will be published open-access. Nonetheless, small additions were made.
Please see our responses to your technical comments below.

> The introduction gives a nice, relatively long (in relation to other
> chapters) overview on different aspects of glacial erosion and resulting
> features, and guides through problems to investigate them. Here, I don’t see
> very clear how different aspects or parts in the introduction are
> specifically addressed in this study, i.e. which problems are exactly aimed
> to be solved of this large portfolio of mentioned gaps in the understanding
> of glacial processes. I think this could be better formulated and balanced.

Our introduction is meant to revolve around two discussion points: the variable
imprint of glaciers on mountain topography, and the link between climate and
glacier erosion (somewhat adressed in the conclusions). This is followed by a
short methodological bibliography of process-based and empirical glacier
erosion laws, before we announce the scope of our work. A few key statements
were added to try and better tie these different sections.

### 2021-05-27 articles/erosion/alps-erosion.tex
Try to improve introduction somewhat.

    The glacial erosion landscape of the Alps has fascinated generations of
    explorers, artists, mountaineers and scientists for centuries. Its cultural
    impact is indeed so far-reaching, that in English, a non-Alpine language,
    the adjective ``alpine'' with non-capital ``a'' is now casually used to
    describe an Alpine-like, glacially modified mountain landscape outside the
    Alps, while the proper noun ``Alps'' has been applied to nick-name
    Alpine-like, glacier eroded mountain ranges in Norway (\emph{Lyngsalpene}),
    New Zealand (Southern Alps), Japan
    (\emph{\begin{CJK}{UTF8}{min}日本アルプス\end{CJK},
    Nihon Arupusu}), and elsewhere.
    %
    Some mountain ranges are predominantly characterised by cirque glaciation
    (e.g. Uinta Mountains), glacial valleys (e.g. Putorana Plateau), or
    large-scale overdeepenings (e.g. Patagonia). But other regions, including
    the Alps \citep[e.g.][Fig.~\ref{fig:landscape}]{Penck.1905}, present a
    higher variety of glacier erosional landforms, whose implications on
    glacial history are yet to be understood.
    The cosmogenic nuclide memory of bedrock erosion provides a more
    quantitative, but equally varied picture of glacier erosion effectiveness
    both within and between glaciated regions \citep{Jansen.etal.2019,
    Steinemann.etal.2020, Steinemann.etal.2021}.

    In the absence of other landforms, glacially eroded topography has
    sometimes been used as a proxy for glacier extent, for instance in
    geomorphological mapping \citep[e.g.,][]{Margold.etal.2011, Fu.etal.2012}.
    Cold-based glaciers, however, have been observed to preserve landforms as
    fragile as sand beaches \citep{Kleman.1994}, leading to the
    reinterpretation of complex glacial landscapes within a binary conceptual
    framework of cold and temperate basal thermal regimes
    \citep{Kleman.etal.2008, Kleman.etal.2010, Fabel.etal.2012, Fu.etal.2013}.
    Nevertheless, it remains unclear how sharp the erosional
    transition to cold-based ice is, and why some glaciated regions show
    preserved periglacial blockfields topped by erratic boulders and others,
    including the Alps, do not \citep{Wirsig.etal.2016, Seguinot.etal.2018}.

    Glacier erosion also likely governs, at least in part,
    the height of mountains \citep{Egholm.etal.2009,Thomson.etal.2010} through
    the ``glacier buzz-saw''. In turn, evidence suggests that increased
    mountain erosion rates coincided with global cooling and Pleistocene
    glaciations from roughly 2.6 million years before present
    \citep{Herman.Champagnac.2016}. However, oceanic isotopic proxies for
    global weathering rates remained constant over this time period
    \citep{Willenbring.Von-Blanckenburg.2010}. Examination of contemporary
    erosion rates across different climates suggests that increased temperatures
    in glaciated regions lead to higher erosion rates and that erosion rates
    may have increased towards present \citep{Koppes.Montgomery.2009,
    Koppes.etal.2015, Fernandez.etal.2016}. However, understanding the link
    between climate and glacier erosion in the sedimentary record remains
    difficult due to time scale biases \citep{Ganti.etal.2016} and the
    non-linear relationship  between atmospheric temperature and erosion
    \citep[e.g.,][]{Anderson.etal.2012,Mariotti.etal.2021}.
    \todo{shorten this paragraph.}

@@

    Here, the non-linear erosion law by \citet{Koppes.etal.2015} is applied to
    previously published model results \citep{Seguinot.etal.2018} and the
    patterns of modelled erosion potential and cumulative last glacial cycle
    erosion
    potential are analysed. We examine the glaciological conditions that cause
    erosion and discuss the implications of these conditions on understanding
    the relationship between climate and glacier erosion. Despite aggregated
    uncertainties on paleoclimate, glacier flow and glacier erosion processes,
    out results provide insights into the diversity of the Alpine glacial
    erosion landscape.

>>>>

    The glacial erosion landscape of the Alps has fascinated generations of
    explorers, artists, mountaineers and scientists for centuries. Its cultural
    impact is indeed so far-reaching, that in English, a non-Alpine language,
    the adjective ``alpine'' with non-capital ``a'' is now casually used to
    describe an Alpine-like, glacially modified mountain landscape outside the
    Alps, while the proper noun ``Alps'' has been applied to nick-name
    Alpine-like, glacier eroded mountain ranges in Norway (\emph{Lyngsalpene}),
    New Zealand (Southern Alps), Japan
    (\emph{\begin{CJK}{UTF8}{min}日本アルプス\end{CJK},
    Nihon Arupusu}), and elsewhere.
    %
    Some mountain ranges are predominantly characterised by cirque glaciation
    (e.g. Uinta Mountains), glacial valleys (e.g. Putorana Plateau), or
    large-scale overdeepenings (e.g. Patagonia). But other regions, including
    the Alps \citep[e.g.][Fig.~\ref{fig:landscape}]{Penck.1905}, present a
    higher variety of glacier erosional landforms, whose implications on
    glacial history are yet to be understood. In parallel to such landscape
    diversity, the cosmogenic nuclide memory of bedrock erosion provides a more
    quantitative, but equally varied picture of glacier erosion effectiveness
    both within and between glaciated regions \citep{Jansen.etal.2019,
    Steinemann.etal.2020, Steinemann.etal.2021}.

    Glacially eroded topography has sometimes been used as a proxy for mapping
    paleoglacier extent \citep[e.g.,][]{Margold.etal.2011, Fu.etal.2012}.
    Cold-based glaciers, however, have been observed to preserve landforms as
    fragile as sand beaches \citep{Kleman.1994}, leading to the
    reinterpretation of complex glacial landscapes within a binary conceptual
    framework of cold and temperate basal thermal regimes
    \citep{Kleman.etal.2008, Kleman.etal.2010, Fabel.etal.2012, Fu.etal.2013}.
    Nevertheless, it remains elusive whether glacial topography is a proxy for
    for temperate ice-cover duration, and why some glaciated regions show
    preserved periglacial blockfields topped by erratic boulders and others,
    including the Alps, do not \citep{Wirsig.etal.2016, Seguinot.etal.2018}.

    On a larger scale, glacier erosion also likely governs, at least in part,
    the height of mountains \citep{Egholm.etal.2009,Thomson.etal.2010} through
    the ``glacier buzz-saw''. In turn, evidence suggests that increased
    mountain erosion rates coincided with global cooling and Pleistocene
    glaciations from roughly 2.6 million years before present
    \citep{Herman.Champagnac.2016}. However, oceanic isotopic proxies for
    global weathering rates remained constant over this time period
    \citep{Willenbring.Von-Blanckenburg.2010}. Examination of contemporary
    erosion rates across different climates suggests that increased temperatures
    within glaciated regions lead to higher erosion rates
    \citep{Koppes.Montgomery.2009, Koppes.etal.2015, Fernandez.etal.2016}.
    However, understanding the link
    between climate and glacier erosion in the sedimentary record remains
    difficult due to time scale biases \citep{Ganti.etal.2016} and the
    non-linear relationship between atmospheric temperature and erosion
    \citep[e.g.,][]{Anderson.etal.2012, Mariotti.etal.2021}.
    Numerical glacier modelling may provide insights into such complexity.

@@

    Here, the non-linear erosion law by \citet{Koppes.etal.2015} is applied to
    previously published model results \citep{Seguinot.etal.2018} and the
    patterns of modelled erosion potential and cumulative last glacial cycle
    erosion potential are analysed. We examine the glaciological conditions
    that cause erosion and discuss the implications of these conditions on
    understanding the mark of ancient glaciers on Alpine topography and
    the relationship between climate and glacier erosion. Despite aggregated
    uncertainties on paleoclimate, glacier flow and glacier erosion processes,
    out results provide insights into the diversity of the Alpine glacial
    erosion landscape.
